Sketch Recognition and Understanding

### 2009-10-10 trunk/draft/Results.tex
after changes of dr. nevin

%For classifier and algorithm testing the system used a data set collected by
\begin{description}
	\item [Hse Dataset] (denoted as Hs-DB) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmark for sketch recognition problem. The data are drawn by 16 users each of them have drawn from 50 to 60 samples for each shape. Figure \ref{fig:symbolSet} shows the shapes used in the data set. The dataset was divided into 75\% training set and 25\% test set. Five different splits for the training and test data are generated from the dataset.% In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respectively. %, 94.4\%  , which
%is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
		\item [Multi Domain Dataset] (denoted as MD-DB) A dataset we collected manually from different users. The dataset contains symbols from two different domains Electrical Engineering symbols (EL-DB) and logic design symbols (LD-DB). Data are collected from 7 different users each of them have drawn from 20 to 40 samples of each shapes. Figures \ref{fig:LogicSet} and \ref{fig:ElectSet} show some of the shapes used in the dataset. Symbols from different domains were gathered to evaluate the system throughout different types of shapes. Users were requested to draw symbols without any restriction on style, size or orientations. The dataset was divided into 75\% training set and 25\% test set. Five different splits for the training and test data are generated from the dataset. %In worth noting that \cite{Oltmans07} gathered similar dataset and achieved 89\% on it   %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols. 
\end{description}

 %We divided the dataset equally into training set and test set.  The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. \\
%We divided the dataset into four sets each set contains four different users. Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.  The algorithms are then tested using the four sets and the results are averaged.
\begin{figure}[]\centering
\fbox{  \parbox{8cm}{% 
		\includegraphics[scale=0.7]{images/symbolSet.eps}	}
		}
	\caption[Hs-DB Symbol Set]{The symboles Set in \cite{HeloiseBeautification} } 
	\label{fig:symbolSet}
\end{figure}

@@

\subsection{Segmentation Algorithms}
\label{sec:SegmentationAlgorithms}
The result of the segmentation algorithm can be viewed in the Figure \ref{fig:results1}. This result shows the originals stroke and the segmentation that the system generated for the stroke. Figure \ref{fig:results2} shows the result of \textsl{AlgS1} and \textsl{AlgS2} for the same input strokes.  Figure \ref{fig:results3} shows the output of \textsl{Alg3} to the same set of input strokes. It clearly shows that \textsl{AlgS2} have better segmentation results than both \textsl{AlgS1} and \textsl{Alg3} but since these segmentation are subjective to what the user intended to draw the next section will focus on the evaluating the segmentation algorithms using the effect on the recognition accuracy.

@@

several experiments is presented to test the different segmentation algorithms on the three data set used in our system. Algorithms (\textsl{AlgS1}, \textsl{AlgS2} and \textsl{Alg3} (in section \ref{sec:BenchMarckAlgorithm})) were tested individually with and without the ellipse fitting module. The final system (Choosing the Minimum error of (\textsl{AlgS1}, \textsl{AlgS2}) was evaluated on the three dataset(HS, EL and LD datasets ). Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} shows the accuracy achieved by each algorithm. Theses experiments were done using the whole dataset as 75\% of the data as test set and 25\% as training. The results shows that \textsl{AlgS2} with the ellipse fitting module gives better recognition rate across all datasets (Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} ). Even though, \textsl{AlgS1} gives better result than \textsl{Alg3} in both EL-DB and HS-DB it perform poorly in LD-DB due to the high number of curves in every symbol in this dataset. These result are proved using different number of samples per category as shown in Figures \ref{fig:LDRes}, \ref{fig:ElRes}. 
% i need to organize this as following. 
% number 1. differnt algorithm on different data 
% which algortihm is better oon which data 
% on ld and el  the alg2  is better,

%We performed several experiments to evaluate the presented recognition system. Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3} in section \ref{sec:BenchMarckAlgorithm} )  to use as reference to compare it with our swarm algorithms. Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD}  shows the accuracy achieved by each algorithm. The two swarm algorithms were tested with and without the ellipse fitting module. The ellipse fitting module appears to be superior. The result shows that combining both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) outperform any single algorithm. %show that combining algorithms achieve higher accuracy than any single algoirthm. 
%improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms. 
\begin{figure}
	\centering		
	 \includegraphics[scale=0.4]{paperImages/testAlg.jpg.eps}
	 	\caption{Algorithm comparison on HS-DB} The recognition rate of different algorithms. 
	 	\label{fig:test1}
	%\caption{Experiments results :} a)    b) c) 
\end{figure} 
\begin{figure}
	\centering		
	 \includegraphics[scale=0.8]{results/AlgLD.eps}
	 	\caption{Algorithm comparison on LD-DB} The recognition rate of different algorithms. 
	 	\label{fig:testLD}
	%\caption{Experiments results :} a)    b) c) 
\end{figure}

\begin{figure}
	\centering		
	 \includegraphics[scale=0.8]{results/AlgEL.eps}
	 	\caption{Algorithm comparison on EL-DB} The recognition rate of different algorithms. 
	 	\label{fig:testEL}
	%\caption{Experiments results :} a)    b) c) 
\end{figure}

@@

 The effects of symbol complexity in a dataset were testing in the following experiment. These experiments test how many samples will be required to finally obtain good performance from the recognizer system. In the following experiments, we tested the system using different number of sample per category. 
 Figure \ref{fig:Alg1Ds} and \ref{fig:Alg2Ds} shows that Hs-DB requires more number of sample for training to achieve same result as in other data sets. This is due that it contains less signal and conceptual variation that the other two datasets. Figure \ref{fig:samplesofHS-DB.jpg} shows sample of data collected from one user in Hs-Db it is clear that it has reasonable amount of variation between users but it is limited. Recognizing symbols that contains high variation in drawing style using fewer symbols in training prove the efficiency of our system.
 \begin{figure}
	\centering
		\includegraphics{results/Alg1Ds.eps}
	\caption{AlgS1 on different datasets}
	\label{fig:Alg1Ds}
\end{figure}
\begin{figure}
	\centering
		\includegraphics{results/Alg2Ds.eps}
	\caption{AlgS2 on different Datasets }
	\label{fig:Alg2Ds}
\end{figure}
 \begin{figure}
	\centering
	\includegraphics[scale=0.5]{results/samplesofHS-DB.jpg.eps}
	\caption[Hs-DB shapes drawn by single user]{Hs-DB shapes drawn by single user \cite{HeloiseBeautification} }
	\label{fig:samplesofHS-DB.jpg}
\end{figure}
  Figure \ref{fig:correct2.jpg} shows examples of correctly recognized symbol using our system. The figure shows that system recognized some hard to recognize example in both EL-DB and DL-DB. An example of symbols that was not correctly recognized are in Figure \ref{fig:erros}. Most of the wrong classifications were due to how messy the user draw symbols and how similarly they are conceptual. Table \ref{tab:ConfusionMatrix} (confusion matrix) shows the types of error in Hs-DB which shows that the system was confused between parallelogram and trapezoid, moon and arch. Those errors were due to users drawing styles as some user may draw a parallelogram as a trapezoid.

@@

 \begin{table*}
 \caption{Confusion Matrix of HS-DB}
	\centering
	%\small
%	\scalebox{0.7}{
		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
	\scalebox{0.5}{
		\begin{tabular}{|p{2cm}|c|c|c|c|c|c|c|c|c|c|c|c|c||p{2cm}|c|c|c|c|p{2cm}|}\hline 
		Categorie / classified as ==$>$	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout	&Number of samples per class	&TP count	&FP count	&Fn	&Precision	&Recall ( TP rate )\\ \hline 
ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0	&152	&150	&27	&2	&0.8474	&0.9868\\ \hline 
heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0	&152	&130	&20	&22	&0.86666	&0.8552\\ \hline 
trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0	&150	&147	&50	&3	&0.7461	&0.98\\ \hline 
pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0	&152	&139	&9	&13	&0.9391	&0.91447\\ \hline 
arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1	&151	&125	&25	&26	&0.83333	&0.82781\\ \hline 
hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0	&151	&135	&2	&16	&0.9854	&0.89403\\ \hline 
square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0	&151	&138	&1	&13	&0.9928	&0.91390\\ \hline 
triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0	&151	&148	&20	&3	&0.88095	&0.98013\\ \hline 
cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0	&153	&146	&6	&7	&0.96052	&0.95424\\ \hline  
cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0	&152	&147	&2	&5	&0.98657	&0.967105\\ \hline 
parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0	&150	&107	&4	&43	&0.963	&0.71333\\  \hline 
moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5	&153	&125	&21	&28	&0.8561	&0.8169\\ \hline  
callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145	&151	&145	&0	&6	&1	&0.9602\\  \hline  
% Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline 
%ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0\\  \hline 
%heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0\\  \hline 
%trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0\\  \hline 
%pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0\\   \hline 
%arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1\\   \hline 
%hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0\\  \hline 
%square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0\\   \hline 
%triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0\\  \hline 
%cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0\\   \hline 
%cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0\\  \hline 
%parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0\\  \hline 
%moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5\\   \hline 
%callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145\\  \hline 
 		\end{tabular}
			}
 		%\end{tabular}
 	%}

@@

Another experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate. Figure \ref{fig:test2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}). It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.  The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines. This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself. Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result. This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation. Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols. The penalty for this improved performance is the computational time required to run both swarm algorithms. Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols. The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles. This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them. This led us to the next experiment to choose best set of feature to recognize symbols.

>>>>

%For classifier and algorithm testing the system used a data set collected by
\begin{description}
	\item [Hse Dataset] (denoted as Hs-DS) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmark for sketch recognition problem. The data are drawn by 16 users each of them have drawn from 50 to 60 samples for each shape. Figure \ref{fig:symbolSet} shows the shapes used in the data set. The dataset was divided into 75\% training set and 25\% test set. Five different splits for the training and test data are generated from the dataset.% In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respectively. %, 94.4\%  , which
%is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
\item [Electrical Dataset] (denoted as EL-DS) A dataset we collected manually from different users. The dataset contains  Electrical Engineering symbols (EL-DS). Data are collected from 7 different users each of them have drawn from 20 to 40 samples of each shapes. Figure \ref{fig:ElectSet} shows some of the shapes used in the dataset. Symbols were gathered to evaluate the system throughout different types of shapes. Users were requested to draw symbols without any restriction on style, size or orientations. The dataset was divided into 75\% training set and 25\% test set. Five different splits for the training and test data are generated from the dataset. %In worth noting that \cite{Oltmans07} gathered similar dataset and achieved 89\% on it   %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols
\item [Digital Logic Dataset](denoted as LD-DS) Similar to EL-Ds the dataset was collected manually from different users. The dataset contains symbols Logic Design symbols (LD-DS). Data are collected from 7 different users each of them have drawn from 20 to 40 samples of each shapes. Figures \ref{fig:LogicSet} shows some of the shapes used in the dataset. The dataset was divided to train and test set similar to EL-DS.  %Symbols from different domains were gathered to evaluate the system throughout different types of shapes. Users were requested to draw symbols without any restriction on style, size or orientations. The dataset was divided into 75\% training set and 25\% test set. Five different splits for the training and test data are generated from the dataset. %In worth noting that \cite{Oltmans07} gathered similar dataset and achieved 89\% on it   %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols
	%	\item [Multi Domain Dataset] (denoted as MD-DB) A dataset we collected manually from different users. The dataset contains symbols from two different domains Electrical Engineering symbols (EL-DB) and logic design symbols (LD-DB). Data are collected from 7 different users each of them have drawn from 20 to 40 samples of each shapes. Figures \ref{fig:LogicSet} and \ref{fig:ElectSet} show some of the shapes used in the dataset. Symbols from different domains were gathered to evaluate the system throughout different types of shapes. Users were requested to draw symbols without any restriction on style, size or orientations. The dataset was divided into 75\% training set and 25\% test set. Five different splits for the training and test data are generated from the dataset. %In worth noting that \cite{Oltmans07} gathered similar dataset and achieved 89\% on it   %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols. 
\end{description}

 %We divided the dataset equally into training set and test set.  The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. \\
%We divided the dataset into four sets each set contains four different users. Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.  The algorithms are then tested using the four sets and the results are averaged.
\begin{figure}[]\centering
\fbox{  \parbox{8cm}{% 
		\includegraphics[scale=0.7]{images/symbolSet.eps}	}
		}
	\caption[Hs-DB Symbol Set]{The symbols Set in \cite{HeloiseBeautification} } 
	\label{fig:symbolSet}
\end{figure}

@@

\subsection{Segmentation Algorithms}
\label{sec:SegmentationAlgorithms}
The result of the segmentation algorithm can be viewed in the Figure \ref{fig:results1}. This result shows the original stroke and the segmentation that the system generated for the stroke. Figure \ref{fig:results2} shows the result of \textsl{AlgS1} and \textsl{AlgS2} for the same input strokes.  Figure \ref{fig:results3} shows the output of \textsl{Alg3} to the same set of input strokes. It clearly shows that \textsl{AlgS2} has better segmentation results than both \textsl{AlgS1} and \textsl{Alg3} but since these segmentations are subjective to what the user intended to draw the next section will focus on evaluating the segmentation algorithms and the effect on the recognition accuracy.

@@

Several experiments are presented to test the different segmentation algorithms on the three data sets used in our system. Algorithms (\textsl{AlgS1}, \textsl{AlgS2} and \textsl{Alg3} (in section \ref{sec:BenchMarckAlgorithm})) were tested individually with and without the ellipse fitting module. The final system (Choosing the Minimum error of (\textsl{AlgS1}, \textsl{AlgS2}) was evaluated on the three dataset(HS, EL and LD datasets ). Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} shows the accuracy achieved by each algorithm. Theses experiments were done using the whole dataset as 75\% of the data as test set and 25\% as training. The results shows that \textsl{AlgS2} with the ellipse fitting module gives better recognition rate across all datasets (Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} ). Even though, \textsl{AlgS1} gives better result than \textsl{Alg3} in both EL-DB and HS-DB it performs poorly in LD-DB due to the high number of curves in every symbol in this dataset. These result are obtained using different number of samples per category as shown in Figures \ref{fig:LDRes}, \ref{fig:ElRes}. 
% i need to organize this as following. 
% number 1. differnt algorithm on different data 
% which algortihm is better oon which data 
% on ld and el  the alg2  is better,

%We performed several experiments to evaluate the presented recognition system. Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3} in section \ref{sec:BenchMarckAlgorithm} )  to use as reference to compare it with our swarm algorithms. Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD}  shows the accuracy achieved by each algorithm. The two swarm algorithms were tested with and without the ellipse fitting module. The ellipse fitting module appears to be superior. The result shows that combining both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) outperform any single algorithm. %show that combining algorithms achieve higher accuracy than any single algoirthm. 
%improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms. 
\begin{figure}
	\centering		
	 \includegraphics[scale=0.4]{paperImages/testAlg.jpg.eps}
	 	\caption{Algorithm Comparison on HS-DB} The recognition rate of different algorithms. 
	 	\label{fig:test1}
	%\caption{Experiments results :} a)    b) c) 
\end{figure} 
\begin{figure}
	\centering		
	 \includegraphics[scale=0.8]{results/AlgLD.eps}
	 	\caption{Algorithm Comparison on LD-DB} The recognition rate of different algorithms. 
	 	\label{fig:testLD}
	%\caption{Experiments results :} a)    b) c) 
\end{figure}

\begin{figure}
	\centering		
	 \includegraphics[scale=0.8]{results/AlgEL.eps}
	 	\caption{Algorithm Comparison on EL-DB} The recognition rate of different algorithms. 
	 	\label{fig:testEL}
	%\caption{Experiments results :} a)    b) c) 
\end{figure}

@@

 The effect of symbol complexity in a dataset is tested in the following experiment. These experiments test how many samples will be required to finally obtain good performance from the recognizer system. In the following experiments, we tested the system using different number of sample per category. 
 Figure \ref{fig:Alg1Ds} and \ref{fig:Alg2Ds} shows that Hs-DB requires more number of samples for training to achieve same result as in other data sets. This is due to the fact that it contains less signal and conceptual variation that the other two datasets. Figure \ref{fig:samplesofHS-DB.jpg} shows sample of data collected from one user in Hs-Db. It is clear that it has reasonable amount of variation between users but it is limited. Recognizing symbols that contains high variation in drawing style using fewer symbols in training prove the efficiency of our system.
 \begin{figure}
	\centering
		\includegraphics{results/Alg1Ds.eps}
	\caption{AlgS1 on Different Datasets}
	\label{fig:Alg1Ds}
\end{figure}
\begin{figure}
	\centering
		\includegraphics{results/Alg2Ds.eps}
	\caption{AlgS2 on Different Datasets }
	\label{fig:Alg2Ds}
\end{figure}
 \begin{figure}
	\centering
	\includegraphics[scale=0.5]{results/samplesofHS-DB.jpg.eps}
	\caption[Hs-DB Shapes Drawn by Single User]{Hs-DB shapes drawn by single user \cite{HeloiseBeautification} }
	\label{fig:samplesofHS-DB.jpg}
\end{figure}

  Figure \ref{fig:correct2.jpg} shows examples of correctly recognized symbol using our system. The figure shows that system recognized some hard to recognize example in both EL-DB and DL-DB. Example of symbols that were not correctly recognized are in Figure \ref{fig:erros}. Most of the wrong classifications were due to how messy the user draws symbols and how similar they are conceptual. Table \ref{tab:ConfusionMatrix} (confusion matrix) shows the types of error in Hs-DB which shows that the system was confused between parallelogram and trapezoid, moon and arch. Those errors were due to users drawing styles as some user may draw a parallelogram as a trapezoid.

@@

 \begin{table*}
 \caption{Confusion Matrix of HS-DB}
	\centering
	%\small
%	\scalebox{0.7}{
		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
	\scalebox{0.5}{
		\begin{tabular}{|p{2cm}|c|c|c|c|c|c|c|c|c|c|c|c|c||p{2cm}|c|c|c|c|p{2cm}|}\hline 
		Category / classified as ==$>$	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout	&Number of samples per class	&TP count	&FP count	&Fn	&Precision	&Recall ( TP rate )\\ \hline 
ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0	&152	&150	&27	&2	&0.8474	&0.9868\\ \hline 
heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0	&152	&130	&20	&22	&0.86666	&0.8552\\ \hline 
trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0	&150	&147	&50	&3	&0.7461	&0.98\\ \hline 
pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0	&152	&139	&9	&13	&0.9391	&0.91447\\ \hline 
arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1	&151	&125	&25	&26	&0.83333	&0.82781\\ \hline 
hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0	&151	&135	&2	&16	&0.9854	&0.89403\\ \hline 
square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0	&151	&138	&1	&13	&0.9928	&0.91390\\ \hline 
triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0	&151	&148	&20	&3	&0.88095	&0.98013\\ \hline 
cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0	&153	&146	&6	&7	&0.96052	&0.95424\\ \hline  
cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0	&152	&147	&2	&5	&0.98657	&0.967105\\ \hline 
parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0	&150	&107	&4	&43	&0.963	&0.71333\\  \hline 
moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5	&153	&125	&21	&28	&0.8561	&0.8169\\ \hline  
callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145	&151	&145	&0	&6	&1	&0.9602\\  \hline  
% Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline 
%ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0\\  \hline 
%heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0\\  \hline 
%trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0\\  \hline 
%pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0\\   \hline 
%arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1\\   \hline 
%hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0\\  \hline 
%square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0\\   \hline 
%triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0\\  \hline 
%cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0\\   \hline 
%cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0\\  \hline 
%parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0\\  \hline 
%moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5\\   \hline 
%callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145\\  \hline 
 		\end{tabular}
			}
 		%\end{tabular}
 	%}

@@

Another experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate. Figure \ref{fig:test2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}). It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols.  The results indicate that algorithm \textsl{AlgS1} achieves better performance than algorithm \textsl{AlgS2} for the symbols that consist only of lines. This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself. Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result. This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation. Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols. The penalty for this improved performance is the computational time required to run both swarm algorithms. Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols. The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles. This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them. This led us to the next experiment to choose best set of feature to recognize symbols.
